Supplemental data to ``\emph{k}-mer-based GWAS enhances causal variant and candidate gene discovery in soybean''

### 2023-02-02 additional_files/additional_file_1.tex
Integrate contents from the supplemental data in the thesis into the supplemental file here

\usepackage[utf8]{inputenc} % declare encoding as utf8
\usepackage{graphicx} % to enable \includegraphics
\usepackage{flafter} % to make sure floats appear after their position in text
\usepackage{pdflscape} % allow some pages to be in landscape orientation
\usepackage{layout} % for debugging purposes, looking at page layout
\usepackage[margin=1in]{geometry} % leaves more space for text body
\usepackage[hidelinks]{hyperref} % hyperlinks from list of figures/tables
\usepackage[labelfont=bf]{caption} % format captions and make links go to figure instead of caption
\usepackage{csvsimple} % read and display csv files as tables
\usepackage{booktabs} % prettier tables
\usepackage{threeparttable} % easy addition of footnotes below tables
\usepackage{textcomp} % makes the \textdegree command available
\usepackage{tocloft} % formatting the lists of tables and figures
\usepackage{xcolor} % enable text coloring
\usepackage{calc} % makes \widthof available to set the width of the last column
\usepackage{array} % using command \arraybackslash to restore the definition of \\ after \centering
\usepackage{longtable} % allow tables to span more than one page
\usepackage{threeparttablex} % threeparttable for longtable
\usepackage{capt-of} % put caption on next page
\usepackage{listings}
\usepackage{datatool} % allows fetching key-value pairs from a text file

@@

% Defining a command for manhattan plot figures
\newcommand{\manhattanplot}[3]{
\begin{cfigure}
	\includegraphics[width = 6in]{#1_manhattan}
	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
	(a) Platypus (SNPs and indels), (b) vg (SVs), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
	The x-axis shows the position along the reference assembly version 4 of Williams82.
	Each point shows the $-\log_{10} (p)$ associated with a particular marker or \emph{k}-mer.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	Vertical dotted lines indicate the position of signals associated with the trait.
	Documented loci are colored according to whether they were found (green) or not (red) by a particular approach.
	The ``Gm'' prefix has been left out of chromosome names for simplicity.
	}
	\label{#1_manhattan}
\end{cfigure}

@@

% Defining a new command for signal plot figures
\newcommand{\signalplot}[4]{
\begin{cfigure}
	\includegraphics[width = 6in]{#1_signal}
	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zommed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using four genotype datasets:
	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
	Panel (a) shows gene models over the genomic interval.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	#4
	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph and vg) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	}
	\label{#1_signal}
\end{cfigure}

@@

% Defining a new command for gene plot figures
\newcommand{\geneplot}[5]{
\begin{cfigure}
	\includegraphics[width = 6in]{#1_gene}
	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using four genotype datasets:
	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
	#5
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	Panel (a) shows gene models over the plotting interval.
	Exons are represented by rectangles whereas introns are represented by horizontal lines.
	Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
	}
	\label{#1_gene}
\end{cfigure}

@@

% Defining a new command for k-mer plot figures
\newcommand{\kmerplot}[5]{
	\begin{cfigure}
		\includegraphics[width = 6.5in]{#1_kmers}
		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
		{Identification of #5 underlying significant \emph{k}-mers at the #2 gene associated with the #3 locus for #4.
		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
		Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
		The red rectangle identifies the region that is zoomed-in in panel (b).
		(b) Nucleotide sequences of haplotypes observed in at least five samples across the dataset.
		Individual nucleotides are colored according to the $-\log_{10}(p)$ of the most significant \emph{k}-mer overlapping them.
		Dashes indicate gaps in haplotype sequence alignment whereas vertical lines indicate differences in sequence between two haplotypes.
		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes correspond to those shown in panel (b).}
		\label{#1_kmers}
	\end{cfigure}

@@

% Defining a new command for LD plot figures
\newcommand{\ldplot}[3]{
	\begin{cfigure}
		\includegraphics[width = 6.5in]{#1_ld}
		\caption[LD plot of significant \emph{k}-mers for #2]{Pairwise LD among #3 significant \emph{k}-mers identified for #2.
		\emph{k}-mers are sorted along the y-axis according to their putative position along the reference assembly version 4 of Williams82,
		as identified by ``Gm'' chromosome labels.
		Sequences that lack a ``Gm'' label (top of the y-axis) represent unanchored scaffolds.
		\emph{k}-mers are represented in the same order along the x- and y-axis.
		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p)$ of each \emph{k}-mer.}
		\label{#1_ld}
	\end{cfigure}

@@

% Allows underscores in captions but is not a great solution
\catcode`_=12

@@

\maketitle \thispagestyle{empty}

\listoftables

\thispagestyle{empty}
%\clearpage

\vspace{4ex}

\listoffigures

%\thispagestyle{empty}

>>>>

\usepackage[utf8]{inputenc} % declare encoding as utf8
\usepackage{graphicx} % to enable \includegraphics
\usepackage{flafter} % to make sure floats appear after their position in text
\usepackage{pdflscape} % allow some pages to be in landscape orientation
\usepackage{layout} % for debugging purposes, looking at page layout
\usepackage[margin=1in]{geometry} % leaves more space for text body
\usepackage[hidelinks]{hyperref} % hyperlinks from list of figures/tables
\usepackage[labelfont=bf]{caption} % format captions and make links go to figure instead of caption
\usepackage{csvsimple} % read and display csv files as tables
\usepackage{booktabs} % prettier tables
\usepackage{threeparttable} % easy addition of footnotes below tables
\usepackage{textcomp} % makes the \textdegree command available
\usepackage{tocloft} % formatting the lists of tables and figures
\usepackage{xcolor} % enable text coloring
\usepackage{calc} % makes \widthof available to set the width of the last column
\usepackage{array} % using command \arraybackslash to restore the definition of \\ after \centering
\usepackage{longtable} % allow tables to span more than one page
\usepackage{threeparttablex} % threeparttable for longtable
\usepackage{capt-of} % put caption on next page
\usepackage{listings}
\usepackage{datatool} % allows fetching key-value pairs from a text file
\usepackage{natbib}

@@

% Defining a command for manhattan plot figures
\newcommand{\manhattanplot}[3]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_manhattan}
	\caption[Manhattan plots for #2]{Manhattan plots generated from the GWAS analysis of #2 on #3 samples using four genotype datasets :
	(a) Platypus (SNPs and indels), (b) vg (SVs), (c) Paragraph (SVs), (d) \emph{k}-mers presence/absence.
	The x-axis shows the position along the reference assembly version 4 of Williams82.
	Each point shows the $-\log_{10} (p)$ associated with a particular marker or \emph{k}-mer.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	Vertical dotted lines indicate the position of signals associated with the trait.
	Documented loci are colored according to whether they were found (green) or not (red) by a particular approach.
	The ``Gm'' prefix has been left out of chromosome names for simplicity.
	}
	\label{#1_manhattan}
\end{cfigure}

@@

% Defining a new command for signal plot figures
\newcommand{\signalplot}[4]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_signal}
	\caption[Zoomed-in Manhattan plots of the #3 locus]{Zoomed-in Manhattan plots of signals detected by the GWAS analysis of #2 at the #3 locus using four genotype datasets:
	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
	Panel (a) shows gene models over the genomic interval.
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	#4
	Gray shaded rectangles indicate the region delimited by the top 5\% (Paragraph and vg) or top 1\% (Platypus and \emph{k}-mers) associations in the signal region.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	}
	\label{#1_signal}
\end{cfigure}

@@

% Defining a new command for gene plot figures
\newcommand{\geneplot}[5]{
\begin{cfigure}
	\includegraphics[width = 5.8in]{#1_gene}
	\caption[Zoomed-in Manhattan plots of the #3 locus gene]%
	{Zoomed-in Manhattan plots generated from the GWAS analysis of #2 at the #4 associated with the #3 locus using four genotype datasets:
	(b) Platypus (SNPs and indels), (c) vg (SVs), (d) Paragraph (SVs), (e) \emph{k}-mers presence/absence.
	#5
	Horizontal dotted lines indicate the 5\% family-wise error-rate significance threshold determined from a randomization approach.
	In the case of Platypus, blue points denote markers used in the original analysis, whereas red points denote points that had originally been pruned but whose
	\emph{p}-values were computed after signal discovery.
	Panel (a) shows gene models over the plotting interval.
	Exons are represented by rectangles whereas introns are represented by horizontal lines.
	Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
	}
	\label{#1_gene}
\end{cfigure}

@@

% Defining a new command for k-mer plot figures
\newcommand{\kmerplot}[5]{
	\begin{cfigure}
		\includegraphics[width = 5.8in]{#1_kmers}
		\caption[Variants associated with significant \emph{k}-mers at the #3 locus]%
		{Identification of #5 underlying significant \emph{k}-mers at the #2 gene associated with the #3 locus for #4.
		(a) Gene model of #2. Exons are represented by rectangles whereas introns are represented by horizontal lines.
		Coding sequences are shown in blue and the direction of transcription is indicated by arrows.
		The red rectangle identifies the region that is zoomed-in in panel (b).
		(b) Nucleotide sequences of haplotypes observed in at least five samples across the dataset.
		Individual nucleotides are colored according to the $-\log_{10}(p)$ of the most significant \emph{k}-mer overlapping them.
		Dashes indicate gaps in haplotype sequence alignment whereas vertical lines indicate differences in sequence between two haplotypes.
		(c) Contingency table of the phenotypes and haplotypes observed in the dataset. Haplotypes correspond to those shown in panel (b).}
		\label{#1_kmers}
	\end{cfigure}

@@

% Defining a new command for LD plot figures
\newcommand{\ldplot}[3]{
	\begin{cfigure}
		\includegraphics[width = 5.8in]{#1_ld}
		\caption[LD plot of significant \emph{k}-mers for #2]{Pairwise LD among #3 significant \emph{k}-mers identified for #2.
		\emph{k}-mers are sorted along the y-axis according to their putative position along the reference assembly version 4 of Williams82,
		as identified by ``Gm'' chromosome labels.
		Sequences that lack a ``Gm'' label (top of the y-axis) represent unanchored scaffolds.
		\emph{k}-mers are represented in the same order along the x- and y-axis.
		The colored rectangles drawn below the x-axis represent the $-\log_{10} (p)$ of each \emph{k}-mer.}
		\label{#1_ld}
	\end{cfigure}

@@

@@

% Detailed methods for some aspects of the paper
\maketitle \thispagestyle{empty}

\listoftables
\listoffigures

\thispagestyle{empty}

\section{Detailed methods}
\label{annexe-sv-gwas-detailed-methods}

\subsection{Discovery of structural variation from Illumina data}
\label{sv-gwas-sv-discovery-illumina}

We used the Illumina WGS data of the 389 selected \emph{G. max} accessions as a
first source of data for SV discovery using methods similar to those used by
\cite{lemay2022}.  Briefly, we used four different SV discovery programs or
pipelines:

\begin{enumerate}

\item We used AsmVar \citep[version of 2015-04-16, ][]{liu2015} to call SVs
	from \emph{de novo} assemblies generated with SOAPdenovo2 v. 2.04
		\citep{luo2012} and aligned to the reference genome using LAST
		v. 1047 \citep{kielbasa2011}.
\item We ran Manta v. 1.6.0 \citep{chen2016} in 78 randomly selected batches of
	5 samples and 1 batch of 4 samples. We used the candidate SVs
		identified from each of the batches for downstream analyses.
\item We used smoove v. 0.2.4 \citep{pedersen2019} with the sequence of
	commands \texttt{smoove call}, \texttt{smoove merge}, \texttt{smoove
		genotype} and \texttt{smoove paste} to obtain a dataset of
		candidate SVs. For three (3) of the 389 samples, we were unable
		to obtain candidate variants from smoove due to a reproducible
		cryptic error (segmentation fault) when running \texttt{smoove
		call}. We nevertheless decided to use smoove given that this
		represented a very small fraction of the total number of
		samples and that other programs were able to successfully
		discover SVs from those samples.
\item We used SvABA v. 1.1.3 \citep{wala2018} to call SVs from each of the
	samples separately. We classified the SVs called as breakends by SvABA
		as either deletions, duplications or inversions, and converted
		them to explicit sequence using our own R script. Variants
		called as indels by SvABA were used as is for downstream
		analyses.
\end{enumerate}

We filtered the SVs called by all tools to remove those smaller than 50 bp or
larger than 500 kb in size, those located on unanchored scaffolds or organellar
genomes, those classified as unresolved breakends (\texttt{SVTYPE=BND}), and
any variant containing at least one N in either the reference or alternate
allele sequence.  We converted all variants to a sequence-explicit format and
normalized them using bcftools \texttt{norm} v. 1.10.2-105
\citep{li2009-samtools}.

We generated a single VCF file from the variants called using Illumina data by
running SVmerge \citep{wong2010} in two steps. In the first step, we merged
variants among the ones called by a given tool by running SVmerge using a
sliding window of 5 bp to merge together very similar variants called by the
same tool. In the second step, we used the output of the first merging steps
for all four tools and merged similar variants across these four datasets using
a sliding window of 15 bp in SVmerge.  We proceeded this way because using a
sliding window of 15 bp on the outputs of all tools without pre-merging
resulted in a computationally intractable number of pairwise comparisons.

\subsection{Discovery of structural variation from Oxford Nanopore sequencing data}
\label{sv-gwas-sv-discovery-ont}

We used SVs discovered among 17 Canadian soybean cultivars sequenced using
Oxford Nano\-pore Technologies (ONT) by \cite{lemay2022} as a second source of
data for SV discovery.  Although that set of cultivars does not overlap the
ones analyzed in this study, the high quality of the SVs that can be discovered
from long-read sequencing data motivated the inclusion of this dataset among
the SV candidates subjected to genotyping.  We merged the SVs detected in the
long-read (ONT) data with the ones discovered in the short-read (Illumina) data
using SVmerge and systematically favored SVs discovered by Illumina whenever
SVs from the two sets were to be merged.  Our rationale for favoring variants
discovered from the Illumina data was that whenever both technologies discover
the same variants, Illumina data is likely to be more precise given its higher
basecalling accuracy.

\subsection{Discovery of structural variation from high-quality assemblies}
\label{sv-gwas-sv-discovery-assembly}

In addition to using the Illumina data from 389 accessions, we also used
high-quality genome assemblies of 29 \emph{G. max} or \emph{Glycine soja}
accessions and called SVs by comparing their sequence to that of Williams82. We
retrieved 26 high-quality genomes assembled by \cite{liu2020-pangenome} as well
as that of the cultivar Zhonghuang 13 (ZH13) from the Genome Warehouse
repository (\url{https://ngdc.cncb.ac.cn/gwh/}).  We also retrieved the genomes
of \emph{G. soja} accession W05 and \emph{G. max} cultivar Lee from SoyBase
\citep{grant2010}.

We used methods based on \cite{liu2020-pangenome} to call SVs based on those
genome assemblies.  First, we used the nucmer program of the MUMmer suite of
tools v. 4.0.0rc1 \citep{marcais2018} with option \texttt{-c 1000} to align
each of the assemblies to the Williams82 assembly. Next, we used the
delta-filter program of the MUMmer suite to keep only one-to-one alignments
between any genome and the Williams82 reference using the \texttt{-1} option.
We used these filtered alignments to call SVs using a version of the svmu
program \citep{chakraborty2019} that we forked from the original repository
(\url{https://github.com/malemay/svmu}) in order to improve execution time and
memory usage.

We used a set of tools that we assembled in an R package called svmutools
\url{https://github.com/malemay/svmutools} to convert the svmu output to VCF
format. We limited ourselves to processing deletions and insertions because of
the complexity of resolving other types of variants from whole-genome
assemblies.  Copy number variants and inversions called by svmu were therefore
excluded from downstream analyses, although we did take inverted regions into
account when determining the sequence of insertions in these regions to ensure
their proper representation in the output VCF file. We also filtered the
resulting set of variants to remove those smaller than 50 bp or larger than 500
kb and those with any N in either the reference or alternate allele sequence.
Finally, we normalized the VCF files using \texttt{bcftools norm} and merged
them using SVmerge to obtain a single VCF file of variants discovered through
comparison of high-quality genome assemblies.

\subsection{Genotyping structural variation using Paragraph}
\label{sv-gwas-sv-genotyping-paragraph}

We used Paragraph v. 2.4a \citep{chen2019} as one of two SV genotyping programs
to generate a set of SV genotypes for the population of 389 samples for use in
GWAS analyses.  The set of SVs used as input for Paragraph was the result of
merging the set of variants discovered from the Illumina and ONT data with
those discovered from the high-quality assemblies and svmu using SVmerge.
Variants in this candidate set were prepared for input to Paragraph by removing
those closer than 150 bp (the maximum read length in our dataset) to any
chromosome end and padding the variant representation using a script written by
\cite{hickey2020} for that purpose.

We used the recommendations by the authors of Paragraph for running their
program at population scale by running Paragraph individually on each sample.
We computed average sequencing depth for each sample from the output of the
\texttt{samtools coverage} command and used 20 times that value for the
\texttt{-M} parameter of Paragraph.  We ran Paragraph using the
\texttt{multigrmpy} command and merged the resulting VCF files using
\texttt{bcftools merge} in order to get a  single genotype file for the whole
population. This VCF was further filtered by setting the genotype calls made
from less than two reads to missing and by removing variants with minor allele
frequency < 0.02 or a proportion of missing data > 0.5. The resulting dataset
of 186,306 SVs was used for GWAS.

\subsection{Genotyping structural variation using the vg toolkit}
\label{sv-gwas-sv-genotyping-vg}

We also used the vg toolkit v. 1.41.0 \citep{hickey2020} and its Giraffe
aligner \citep{siren2021} for genotyping our set of candidate SVs on the whole
population. The reason for also using vg is that its variation graph model allows
the integration of SNPs and indels and makes for a more robust and integrated
approach to SV genotyping, even though previous benchmarks comparing those two
programs in soybean showed a reduced sensitivity of the vg toolkit in
genotyping insertions \citep{lemay2022}.

We constructed the variation graph using the \texttt{vg autoindex} command with
the \texttt{-\-workflow giraffe} option. We used the Williams82 reference
assembly version 4 as a basis for the graph, to which we added the set of
candidate SVs described above and the set of 7M SNPs and indels called by
Platypus.  We aligned the trimmed reads of the 741 SRA runs in FASTQ format to
the variation graph using the \texttt{vg giraffe} command with default
parameters. Reads that were still paired after adapter- and quality- trimming
were aligned in paired-end mode, whereas the remaining reads were aligned in
single-end mode. The resulting files in GAM format were used to compute read
coverage on the graph using the \texttt{vg pack} command with options
\texttt{-Q5 -s5}. GAM files belonging to the same accession but originating
from different SRA runs were concatenated to be used as input for the
\texttt{vg pack} command such that a single coverage file per accession was
obtained as output of this step. The command \texttt{vg call} was then used to
call genotypes based on read coverage using the parameters
\texttt{-\-genotype-snarls -\-min-length 50} in order to call all variants
$\geq$ 50 bp supported by the graph.

We merged the genotypes called on all samples using \texttt{bcftools merge} to
obtain a single VCF file for the population. Multiallelic records were split
into biallelic records after merging. Genotype calls made with a genotype
quality (GQ) value < 10 were set to missing, and variants were filtered for a
fraction of missing genotypes $\leq$ 0.5 and a minor allele frequency $\geq$
0.02. The resulting VCF file, which was used for GWAS, contained 51,960 SVs.

\subsection{Comparison of WGS and SoySNP50K genotypes}
\label{annexe-sv-gwas-soysnp50k}
When we first analyzed the results from the GWAS analysis on all 389 samples,
we found discrepancies between the haplotypes at some loci and the expected
phenotype that should correspond to that haplotype for simple traits such as
flower color and seed coat color. Upon further analysis, we found that a
mismatch between the sequencing data and the identity of the accession was the
most probable cause of these discrepancies. In order to filter out all samples
for which this could be the case, we compared the genotypes of SNPs obtained
from the SoySNP50K array \cite{song2013} to those obtained from the Illumina WGS
data of the 385 samples that had a match in the GRIN database. Briefly, we
retrieved SoySNP50K genotype calls from SoyBase \citep{grant2010} and extracted
all SNPs with a MAF $\geq$ 0.1 in our population. We translated the positions
of these SNPs from assembly version 2 to assembly version 4 of Williams by
finding exact matches of 41-bp sequences surrounding the SNP positions. The
32,852 SNPs for which unambiguous positions could be found on genome assembly
version 4 were genotyped from the mapped WGS reads using \texttt{bcftools
mpileup} and \texttt{bcftools call}. We then used \texttt{bcftools gtcheck} to
compute the concordance of genotype calls between WGS samples and their
expected match in the SoySNP50K genotype dataset.  We identified 24 samples
with < 90\% concordance between the two genotype datasets and removed them from
the dataset, in addition to another sample for which concordance was over 90\%
but poor raw sequencing data quality (\%GC content) suggested contamination.
Following these filtering steps, 363 samples remained for GWAS analyses. The
concordance between WGS data and SoySNP50K genotypes is included in
Additional file 2.

% Detailed results that are not meant to fit the main paper
\section{Detailed results}
\label{annexe-sv-gwas-detailed-results}

This section provides a more detailed account of some the the results obtained
by the GWAS analyses on a per-trait basis. Aspects of the results that were
already discussed in the main text are not repeated here; in such cases,
readers are simply redirected to specific tables and figures containing
detailed results from all four GWAS approaches. A summary of the signals found
at all loci using all four methods is presented in Table \ref{signals_table}.
Readers are invited to refer to this table as we go over the results obtained
for each trait below.

\subsection{Flower color}
\label{sv-gwas-flower-color}

We conducted a single GWAS analysis contrasting purple and white flower colors
(Table \ref{FLOWER.COLOR_gwas_table}). Of the two classical loci reported by
\cite{bandillo2017}, \textit{W1} was detected by all four GWAS approaches whereas
\textit{L1} was detected by none of the approaches (Figure
\ref{flower_color_manhattan}). These results are unsurprising as W1 is the
major contributor to flower color in soybean, whereas the \textit{L1} locus
(typically associated with pod color) was detected with a much less significant
\textit{p}-value (see Table \ref{signals_table}) and only conjectured as
associated with flower color by \cite{bandillo2017}. Detailed results for the
\textit{W1} locus can be viewed in figures \ref{flower-color-main-figure},
\ref{flower_color_W1_signal} and \ref{flower_color_W1_gene}.

The \textit{k}-mer GWAS analysis also identified presumably spurious significant
\textit{k}-mers scattered throughout the genome (Figure
\ref{flower_color_manhattan}). An analysis of the pairwise LD between a subset
of those \textit{k}-mers showed that all \textit{k}-mers formed a single LD block
linked to the \textit{W1} locus (Figure \ref{flower_color_ld}), ruling out the
presence of a second locus controlling that trait in this population.

\subsection{Pubescence color}
\label{sv-gwas-pubescence-color}

We conducted two GWAS analyses on pubescence color. The first analysis
contrasted gray, tawny and light tawny pubescence colors (Table
\ref{PUBESCENCE.COLOR_gwas_table}) and detected a signal overlapping the
\emph{T} locus using all four approaches (Figure
\ref{pubescence_color_all_manhattan}). Detailed results at the \textit{T} locus
van be viewed in figures \ref{pubescence_color_all_T_signal},
\ref{pubescence_color_all_T_gene} and \ref{pubescence_color_all_T_kmers}. We
observed that not all accessions with gray pubescence harbored the deleted
nucleotide, which could indicate either that other causal variants resulting in
a non-functional F3’H are present in this population, or that the causal
variant lies elsewhere.

This first GWAS analysis also revealed a second locus on chromosome Gm16
(Figure \ref{pubescence_color_all_manhattan}). Upon closer analysis, this locus
appeared to be in strong LD with \emph{k}-mers mapping to chromosome Gm01 and
most importantly in moderate LD with the \emph{T} locus (Figure
\ref{pubescence_color_all_ld}), which suggests that this does not represent an
independent locus controlling pubescence color. This interpretation is also
supported by the fact that the signal at that locus only spans about 10 kb on
chromosome Gm16.

In the presence of a functional allele at the \emph{T} locus, the genotype at
the \emph{Td} locus influences the exact shade of pubescence color.  Therefore,
we conducted a second GWAS analysis contrasting accessions with tawny and light
tawny pubescence in order to ignore the epistatic effect of the \emph{T} locus
and instead focus on the effect of the \emph{Td} locus on pubescence color
(Table \ref{PUBESCENCE.COLOR_gwas_table}). Detailed results at this locus
can be viewed in figures \ref{pubescence-color-main-figure}, \ref{pubescence_color_nogray_Td_signal} and
\ref{pubescence_color_nogray_Td_gene}).

\subsection{Seed coat color}
\label{sv-gwas-seed-coat-color}

We conducted two different GWAS analyses on seed coat color. The first GWAS
included five different phenotypes observed in the dataset (Table
\ref{SEED.COAT.COLOR_gwas_table}) and resulted in the detection of strong
signals overlapping the \emph{I} locus by all four approaches (Figure
\ref{seed_coat_color_all_manhattan}). All approaches but vg also detected
signals at the \emph{G} locus. In addition, the \emph{k}-mer approach also
detected weak signals overlapping the \emph{T} and \emph{O} loci. However, the
\emph{T} and \emph{O} loci are likely chance associations given the low \textit{p}-value
associated with the \emph{T} locus and the proximity of the \emph{O} locus to
the \emph{I} locus. Detailed results of this first GWAS and the \textit{I}
locus can be viewed in figures \ref{seed_coat_color_all_manhattan} and
\ref{seed_coat_color_all_I_signal}.

The second GWAS analysis on seed coat color focused on a subset of accessions
with yellow or green seed coat only, as this difference is controlled by the
\emph{G} locus (Table \ref{SEED.COAT.COLOR_gwas_table}). Detailed results of
this analysis can be viewed in figures \ref{seed-coat-color-main-figure},
\ref{seed_coat_color_greenyellow_manhattan},
\ref{seed_coat_color_greenyellow_G_signal} and
\ref{seed_coat_color_greenyellow_G_gene}.

\subsection{Stem termination type}
\label{sv-gwas-stem-termination}

We conducted two GWAS analyses on stem termination type. The first GWAS
included all phenotypes observed in the dataset (determinate, semi-determinate,
and indeterminate; Table \ref{STEM.TERMINATION.TYPE_gwas_table}). The results
of this analysis and the \textit{Dt1} locus can be viewed in figures \ref{stem_termination_all_manhattan}
\ref{stem_termination_all_Dt1_signal}, \ref{stem_termination_all_Dt1_gene} and
\ref{stem_termination_all_Dt1_kmers}.

The second GWAS analysis that we conducted on this trait included only
semi-determinate and indeterminate phenotypes in an attempt to detect signals
at the \emph{Dt2} locus (Table \ref{STEM.TERMINATION.TYPE_gwas_table}).
However, this was unsuccessful using all four GWAS
approaches (Figure \ref{stem_termination_sn_manhattan}). Similarly, we failed to
detect signals at the \textit{E3} locus, which is more commonly associated
with maturity but was mentioned as associated with stem termination type by
\cite{bandillo2017}.

\subsection{Hilum color}
\label{sv-gwas-hilum-color}

We conducted three different GWAS analyses on hilum color. The first GWAS
included seven colors (Table \ref{HILUM.COLOR_gwas_table}) and detected signals
at the \emph{I}, \emph{T} and \emph{R} loci using all four approaches except vg
at the \emph{R} locus (Figure \ref{hilum_color_all_manhattan}) while no signals
were detected at the \emph{O} and \emph{W1} loci studied by
\cite{bandillo2017}. The detailed results for the \textit{I} locus can be viewed
in figure \ref{hilum_color_all_I_signal} whereas those for the \textit{T} locus
can be viewed in figures \ref{hilum_color_all_T_signal}, \ref{hilum_color_all_T_gene}
and \ref{hilum_color_all_T_kmers}.

Although \cite{bandillo2017} listed the \textit{W1} locus (typically associated
with flower color) as putatively associated with hilum color, we found no
evidence of association in our results.

A second GWAS analysis contrasting accessions with either black or brown hilum
color was better suited at resolving the signal at the \emph{R} locus (Table
\ref{HILUM.COLOR_gwas_table}). Accordingly, this GWAS analysis detected strong
signals overlapping the \emph{R} locus and no other classical loci using all
four approaches (Figure \ref{hilum_color_blackbrown_manhattan}).
\cite{gillman2011} identified the R2R3 MYB transcription factor Glyma.09g235100
as the molecular basis for the \emph{R} locus and documented four different
loss-of-function alleles linked to brown hilum color. Although our analyses did
detect strong signals in the vicinity of this gene, the most associated markers
were not located in the body of the gene (Figure
\ref{hilum_color_blackbrown_R_signal}). Interestingly, the most significantly
associated \emph{k}-mers detected by our analysis mapped 2 kb upstream of
another gene (Glyma.09g234900) annotated as a MYB transcription factor and
putatively involved in the regulation of anthocyanin biosynthesis. This gene
as well as Glyma.09g235100 and two others are part of a tandem array of four
R2R3 MYB genes considered candidates by \cite{gillman2011}, and of which
Glyma.09g235100 was the only one expressed in the seed coat. Given the
convincing evidence provided by \cite{gillman2011}, it is most likely that
Glyma.09g235100 is indeed the gene associated with the \emph{R} locus and that
our failure to identify the most significantly associated \emph{k}-mers as
causal variants is due to the coexistence of several variants that can cause
loss of function at the \emph{R} locus. Indeed, we observed at least two causal
variants (a 1-bp deletion and a SNP) previously documented by
\cite{gillman2011} in our dataset (Figure \ref{hilum_color_blackbrown_R_gene},
\ref{hilum_color_blackbrown_R_kmers}), although they were not among the most
significantly associated markers or \emph{k}-mers.

We conducted a third GWAS focusing on the brown and reddish brown hilum color
phenotypes in order to detect a signal at the \emph{O} locus, however this GWAS
did not detect associations at any of the classical loci (Figure
\ref{hilum_color_rbr_manhattan}).

\subsection{Pod color}
\label{sv-gwas-pod-color}

We conducted two GWAS analyses on pod color. The first analysis included four
different colors (Table \ref{POD.COLOR_gwas_table}) and resulted in the
detection of signals overlapping both the \emph{L1} and \emph{L2} loci with all
four methods (Figure \ref{pod_color_all_manhattan}). Detailed results at the
\textit{L1} locus can be viewed in figure \ref{pod_color_all_L1_signal},
whereas detailed results at the \textit{L2} locus can be viewed in figure
\ref{pod_color_all_L2_signal}.

The second GWAS contrasted black and brown pods with the objective of detecting
signals that provided better resolution at the \emph{L1} locus (Table
\ref{POD.COLOR_gwas_table}).  Accordingly, all four methods detected a signal
at the \emph{L1} locus but not at the \emph{L2} locus using this GWAS (Figure
\ref{pod_color_blbr_manhattan}).  Results at the \emph{L1} locus were similar
to those observed when using all observed phenotypes, but the \emph{k}-mer
analysis revealed a previously undocumented signal on chromosome Gm15 which was
not in LD with any other signal (Figures \ref{pod_color_blbr_manhattan},
\ref{pod_color_blbr_ld}). The observed signal at this locus overlaps a total of
27 genes (Figure \ref{pod_color_blbr_pdcGm15_signal}), but none of those was
annotated as having an obvious role in controlling pod color.

\subsection{Pubescence form}
\label{sv-gwas-pubesence-form}

We conducted two different GWAS on pubescence form. The first GWAS used three
phenotypes (erect, semi-appressed and appressed; Table
\ref{PUBESCENCE.FORM_gwas_table}) and detected signals overlapping the
\emph{Pa1} locus using all four approaches as well as signals overlapping the
\emph{Pa2} locus using Platypus and the \emph{k}-mer approaches (Figure
\ref{pubescence_form_all_manhattan}). Detailed results at the \textit{Pa1} locus
can be viewed in figure \ref{pubescence_form_all_Pa1_signal},
\ref{pubescence_form_all_Pa1_gene} and \ref{pubescence_form_all_Pa1_kmers},
whereas detailed results at the \textit{Pa2} locus can be viewed in figure
\ref{pubescence_form_all_Pa2_signal}.

Signals were also observed on chromosomes Gm04 and Gm15 by the Platypus and
\emph{k}-mer analyses (Figure \ref{pubescence_form_all_manhattan}), however
upon further analysis of LD patterns they appeared to be linked to the
\emph{Pa1} locus (Figure \ref{pubescence_form_all_ld}).

A second GWAS analysis focused only on appressed and semi-appressed phenotypes
with the objective of enhancing the signal at the \emph{Pa2} locus. However,
this analysis only detected weak signals at the \emph{Pa1} locus (Figure
\ref{pubescence_form_noerect_manhattan}) and its results were therefore not
analyzed further.

\subsection{Pubescence density}
\label{sv-gwas-pubescence-density}

We conducted a single GWAS analysis on pubescence density (Table
\ref{PUBESCENCE.DENSITY_gwas_table}), which detected signals at the \emph{Ps}
locus but not the \emph{Pd1} and \emph{P1} loci using all four approaches
(Figure \ref{pubescence_density_manhattan}). \cite{liu2020ps} linked the gene
Glyma.12g187200 to the \emph{Ps} locus and identified copy number variation
(CNV) of a 25.6-kb sequence including this gene as the causal variant, with
higher copy number resulting in decreased pubescence. Through a segmentation
analysis of the sequencing depth in accessions with increased copy number, we
identified the boundaries of the region varying in copy number as 36,254,989 -
36,297,368 on chromosome Gm12 of genome assembly version 4 of Williams82, for a
total of about 42 kb. Detailed results at this locus can be viewed in figures
\ref{pubescence_density_Ps_signal}, \ref{pubescence_density_Ps_gene}, and
\ref{pubescence-density-main-figure}.

\subsection{Seed coat luster}
\label{sv-gwas-seed-coat-luster}
